---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(JTPfunc)
library(dplyr)
library(tibble)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Package JTPfunc



A random collection of miscellaneous functions. 

### Installation

You can install the development version of JTPfunc from [GitHub](https://github.com/) with:

``` r 
# install.packages("devtools")
devtools::install_github("justinpriest/JTPfunc")
```

## Example Functions

There are many unique functions contained here, many with esoteric applications 
to ADF&G data.  


### Function `statweek()`

To calculate the statistical week (Sunday-Saturday), use function `statweek()`
on a date object to return the stat week: 
```{r statweekdemo}
statweek(as.Date("2021-07-01"))
```


### Function `theme_crisp()`

One applicable extension is a ggplot theme called with `theme_crisp()`:
```{r exampleplot, fig.height = 3, fig.width = 7, warning = FALSE, message = FALSE}
library(JTPfunc)
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg, color = wt)) + geom_point() + theme_crisp()
```


### Function `duplicaterows()`

For certain types of date (e.g., age summaries from the scale database)
the data are exported in an already summarized manner. That is,
all fish of the same date, species, sex, length, etc. are combined 
into a single row.  
If we wish to run statistics on these data, it is helpful to have each
row be its own observation.  
Our example data look like: 

```{r cohodataimp, include = FALSE}
coholengthdata <- tibble::tribble(
                    ~Species, ~Sex, ~Length, ~Count,
                      "Coho",  "M",    621L,     1L,
                      "Coho",  "F",    497L,     2L,
                      "Coho",  "M",    557L,     3L,
                      "Coho",  "F",    563L,     1L
                    )


```

```{r cohodatasumm}
coholengthdata 
```

To add new rows to these data, we use function `duplicaterows()`, specifying 
which row is the number of times to repeat. We can see that the row
with the two 497 mm female coho is repeated twice while the row with 
three 557 mm male coho is repeated 3 times:  
```{r cohodatafxn}
duplicaterows(dataframename = coholengthdata, 
              duplicatecolname = "Count")

```



### Function `count_pct()`

A common request is to determine the proportion of observations for each group 
of data. Assuming that each row is an individual, we can calculate a quick 
summary using `count_pct()` like so: 
```{r countpctdemo}
count_pct(iris %>%
           group_by(Species))
```


### Function `addrowconditional()`
For the following example, we'll use hypothetical rockfish catch data.  
In some historical datasets, catches were not identified to a specific species 
but rather to species aggregates. For example, species code 168 refers to 
unspecified demersal shelf rockfishes while species code 140 is red rockfishes. 

```{r loaddata, include = FALSE}
rockfishcatch <- tibble::tribble(
  ~Year, ~Location, ~Species, ~Catch,
  2010L,   "Outer",     140L,    12L,
  2010L,   "Inner",     140L,    33L,
  2010L,  "Middle",     140L,    19L,
  2010L,   "Outer",     168L,    33L,
  2010L,   "Inner",     168L,    27L,
  2010L,  "Middle",     168L,    24L,
  2010L,   "Outer",     110L,    33L,
  2010L,   "Inner",     110L,    27L,
  2010L,  "Middle",     110L,    24L,
  2011L,   "Outer",     140L,    12L,
  2011L,   "Inner",     140L,    33L,
  2011L,  "Middle",     140L,    19L,
  2011L,   "Outer",     168L,    33L,
  2011L,   "Inner",     168L,    27L,
  2011L,  "Middle",     168L,    24L 
)
rockfishcatch <- rockfishcatch %>%
  mutate(Location = as.factor(Location))

```

```{r showdata}
head(rockfishcatch)
```

For these data, the total catch of the species aggregates (species 140 and 168) 
was composed of several species. For this reason, these species may 
require adding additional rows so that the total catch may be apportioned. 
In our example, we repeat all rows of species 168 three times, and 
repeat rows of species 140 two times, then sort by year and species:
```{r addrowconddemo}
addrowconditional(rockfishcatch, criteriacolumn = Species,
                   repeatcount1 = 3, repeatcount2 = 2,
                   criteria1 = 168, criteria2 = 140,
                   sort1 = Year, sort2 = Species)
```



